---
title: "Usage of package pluviometer to estimate the amount of rain using a home-made pluviometer"
author: "Marco Bascietto"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message=FALSE}
library(dplyr)
library(lubridate)
library(pluviometer)
library(ggplot2)
library(tidyr)
```

Let's follow a typical workflow to convert readings from home-made rain gauges (pluviometers) into precipitation figures and to estimate the recharge of a water tank connected to a house roof thanks to precipitation on the roof itself.

Having built a couple of rain gauges with different funnel diameters, let's import a couple of readings we gathered during precipitation events:

```{r}
precipitation_events <- read.table(
  text = "
funnel_diameter_cm water_volume_ml date_dmy
11                 13              20/06/2015
25                 329             24/06/2015
25                 100             02/07/2015
25                 200             03/07/2015
25                 500             16/07/2015",
  header = TRUE)
```


`funnel_diameter_cm` is the diameter of the circular opening where rain is captured, `water_volume_ml` is the amount of water collected in the rain gauge (either as ml or as g).

The surface area of the house roof:
```{r}
roof_area <- 12*12 # m2
discharge_coefficient <- 0.8
```

Now for the core parte: let's compute the amount of precipitation from the rain gauge readings for each precipitation event:

```{r}
precipitation_events <- precipitation_events %>%
  mutate(
    date = lubridate::dmy(date_dmy),
    funnel_area_cm2 = get_funnel_area(funnel_diameter_cm), # cm2
    pluv_factor     = get_pluviometer_factor(funnel_area_cm2), # 1/m2
    rain_mm         = get_precipitation_measure(water_volume_ml, pluv_factor), # ml o l/m2
    tank_rain_charge_l = get_tank_charge(rain_mm, roof_area, discharge_coefficient) # l
  )
print(precipitation_events %>% select(date, rain_mm, tank_rain_charge_l))
```

Finally a monthly summary:

```{r}
monthly_precipitation <- precipitation_events %>%
  mutate(year = year(date), month = month(date, label = TRUE)) %>%
  group_by(month, year) %>%
  select(rain_mm, tank_rain_charge_l) %>%
  summarise_each(funs(sum))

print(monthly_precipitation)
```

## Water balance in the water recycling tank


```{r}
sprinkler_flow <- 10 # l/min
sprinkler_time_min <- c(rep(10, 3), rep(5, 6)) # minutes/day
sprinkler_daily_vol_l <- Reduce(f = sum, sprinkler_time_min * sprinkler_flow) # l/day
tank_volume_l <- 5000 # l
```

A `r tank_volume_l` litre water tank serves an irrigation system that includes `r length(sprinkler_time_min)` sprinklers that need `r sprinkler_flow` litre/min of water each on a daily basis.

Let's now simulate 2 months of lawn irrigation (June and July 2015), eg `r sprinkler_daily_vol_l` litres of water daily:

```{r}
water_use_tbl <- data.frame(
  date                = as.Date("2015/06/01") + 0:60,
  daily_water_usage_l = sprinkler_daily_vol_l)
```

We need to know how much rain water went into the tank in the period June-July 2015. Let's join the lawn irrigation table to the precipitation table:

```{r}
water_use_tbl <- water_use_tbl %>% 
  left_join(
    precipitation_events %>% 
      select(date, tank_rain_charge_l),
    by = "date")
```

When the irrigation season starts (June, 1st) the tank is full of rain water (`r tank_volume_l`) litres:

```{r}
water_use_tbl$water_balance_l    <- NA_real_
water_use_tbl$water_balance_l[1] <- tank_volume_l
water_use_tbl$recharge <- FALSE
water_use_tbl$rain_lost_l <- 0
```

Every time the tank empties it need to be refilled with water from the well. Let's estimate how much water we have pumped from the well, as opposed to the water we recycled from rain fallen on the house roof, on a daily level.

Firstly we need a simulation function that lowers water level in the tank according to sprinkler water needs, pumps water from the well when water level drops below 0, and adds rain water, on a daily level:

```{r}
get_water_balance <- function(water_balance_l, tank_rain_charge_l, daily_water_usage_l, recharge, rain_lost_l, tank_volume_l) {

  new_water_balance_l <- sum(c(
    water_balance_l, 
    tank_rain_charge_l, 
    daily_water_usage_l*-1), 
    na.rm = TRUE)

  if (new_water_balance_l < 0) {
    new_water_balance_l <- new_water_balance_l + tank_volume_l
    recharge  <- TRUE
  }
  if (new_water_balance_l > tank_volume_l) {
    rain_lost_l <- new_water_balance_l - tank_volume_l
    new_water_balance_l   <- tank_volume_l
    tank_rain_charge_l <- tank_rain_charge_l - rain_lost_l
  }
  water_balance_l <- new_water_balance_l
  data.frame(water_balance_l, tank_rain_charge_l, daily_water_usage_l, recharge, rain_lost_l)
}
```

Then we walk the `water_use_tbl` data frame:
```{r}
for(i in seq(1, nrow(water_use_tbl), by = 1)) {
  day_row <- water_use_tbl[i,]

  if(i == 1) {
      prior_water_balance_l <- day_row$water_balance_l
  } else {
      prior_water_balance_l <- water_use_tbl$water_balance_l[i-1]
  } 

  tmp_df <- with(
    day_row, 
    get_water_balance(
      prior_water_balance_l, 
      tank_rain_charge_l, 
      daily_water_usage_l, 
      recharge, 
      rain_lost_l, 
      tank_volume_l))

  water_use_tbl[i,] <- within(
    water_use_tbl[i,], {
      water_balance_l = tmp_df$water_balance_l
      recharge        = tmp_df$recharge
      rain_lost_l     = tmp_df$rain_lost_l
      tank_rain_charge_l = tmp_df$tank_rain_charge_l
  })
}

```


```{r fig.width=7}
water_use_tbl %>% 
  select(date, water_balance_l, rain_lost_l) %>% 
  gather(key = "variable", value = "value", -date) %>% 
  ggplot(., aes(x = date, y = value, color = variable)) + 
    geom_line() + 
    scale_y_continuous("Volume (l)") +
    scale_x_date("Date") +
    ggtitle("Balance in a water recycling tank") +
    theme_light()
```

It looks like a lot of water is pumped from the well, and occasionally rain water is lost due to the tank having been filled just days before. We also need a much bigger water tank to store rain water... 

Finally we can estimate how much water we recycled from rainfall and how much water we pumped from the well:
```{r}
total_water_budget <- water_use_tbl %>%
  mutate(
    tank_well_charge_l = recharge * tank_volume_l) %>% 
  select(
    tank_rain_charge_l, 
    tank_well_charge_l, 
    rain_lost_l) %>% 
  summarise_each(
    funs(sum(., na.rm = TRUE)))
total_water_budget %>% print
```

Hopefully next year won't be as dry as 2015...
